---
title: "Comparison of classification models"
author: "Cesar Conejo"
date: "March 7, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Classification algorithms

In classification algorithms, the purpose is to find a model (a function or something) to predict the class to which each record would belong. Formally, given a database $D$ = {t~1~, t~2~, ..., t~n~} of tuples or records
(individuals) and a set of classes $C$ = {t~1~, t~2~, ..., t~m~}, the classification problem is finding a function $$f: D \rightarrow C $$ such that each t~i~ is assigned a class C~j~.


This function $f$ can be anyone of this models:

- Support Vector Machine

## Modeling - Evaluation Metrics

To determine how good is a model, we have the following metrics

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
